---
title: "Section 4: Post-hoc 1: Association between extreme ferritin levels and breast cancer risk (case-cohort analysis). Restricted to women who are premenopausal at baseline"
output:
  pdf_document:
    toc: no
    number_sections: true
    includes:
      in_header: header.tex 
  html_document:
    theme: united
    toc: yes
geometry: margin=1.5cm
always_allow_html: true
---


# Section 4: Post-hoc 1: Association between extreme ferritin levels and breast cancer HR (case-cohort analysis). Restricted to women who are premenopausal at baseline.


- Define high levels as $\geq$ 200 $\mu$g/L and 300 $\mu$g/L for premenopausal and postmenopausal females, respectively. [@koperdanovaInterpretingRaisedSerum2015].

- Iron deficiency commonly identified as < 30 $\mu$g/L [@soppiIronDeficiencyAnemia2018].
    - Values can range from 12 to 100 $\mu$g/L [@peyrin-birouletGuidelinesDiagnosisTreatment2015]. 
    - WHO definitions have a threshold at < 15 $\mu$g/L. 

```{r, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      results = 'markup',
                      warning = F)

```


```{r, include=F}
# bring in packages
require(knitr)
require(data.table)
require(tableone)
require(ggplot2)
require(kableExtra)
require(rowr)
require(survival)
require(stargazer)
require(splines)
require(ggplot2)
require(survminer)

require(dplyr)
require(reshape2)
```


```{r}

# source: section1.Rmd
load(file="../sections/updated-data-iron.RData") # df3 is the iron subset

df = data.frame(df3)

```

```{r}

summary(df[df$menop.status==0 & df$fu.meno.age<df$start.age,]$fu.meno.age)

head(df[df$menop.status==0 & df$fu.meno.age<df$start.age, c("start.age", "fu.meno.age", "menop.status")])
table(df$menop.status)

table(df$menop.status.f)

```


```{r, results='hide'}

# handle data for survival analysis

df = within(df, {

  
  fe.log = log(UMN_Iron_Baseline_FE)
  fertn.log = log(UMN_Iron_Baseline_FERTN)
  fesat.log = log(UMN_Iron_Baseline_FESAT)
  tibc.log = log(UMN_Iron_Baseline_TIBC)
  uibc.log = log(UMN_Iron_Baseline_UIBC)
    
  fe.scale = scale(log(UMN_Iron_Baseline_FE), scale=F)
  fertn.scale = scale(log(UMN_Iron_Baseline_FERTN), scale=F)
  fesat.scale = scale(log(UMN_Iron_Baseline_FESAT), scale=F)
  tibc.scale = scale(log(UMN_Iron_Baseline_TIBC), scale=F)
  uibc.scale = scale(log(UMN_Iron_Baseline_UIBC), scale=F)

})

# took out 'ever.donate.b.f' 9/12/2019
confounder.list = c('smoke.f2', 'alc.f2', 'educ.f2', 'ever.hrt.f', 'birth.control.f',
                    'one.term.birth.f')

# Note: Still waiting on  multivitamin with minerals (iron)

# time-dept menopause ==============================

summary(df$fu.meno.age) # 892 missing
df$miss.meno.age = with(df, is.na(df$fu.meno.age))
with(df, table(miss.meno.age, menop.status)) # 27 people are post-menopausal at baseline but no age at menopause. Put age at menopause at baseline age.

# should I assume they are still pre-menopausal at baseline?
df$fu.meno.age.rev = with(df, ifelse( is.na(fu.meno.age)==T & menop.status==1, baseline.age,
                                    ifelse(is.na(fu.meno.age)==T & menop.status==0, 99, fu.meno.age )))
summary(df$fu.meno.age.rev)
summary(df[df$menop.status==0,]$fu.meno.age.rev)

df[df$PSID %in% c("00224_100003"), c("fu.meno.age", "baseline.age", "menop.status", "fu.meno.age.rev")]

df$miss.meno.age.rev = with(df, is.na(df$fu.meno.age.rev))
with(df, table(miss.meno.age.rev, menop.status)) # check

```


```{r}
# right censor premenopausal women at baseline at age at menopause if it exists.
# borrowed following code from section2-cch-premenopause.Rmd

df$c.age.alt = with(df, ifelse(menop.status==1, c.age, # if a person is postmenop at baseline then keep original censoring age
                               ifelse(menop.status==0 & c.age>fu.meno.age.rev & fu.meno.age.rev>start.age, fu.meno.age.rev, # premenopausal at baseline and censoring age after menopause age then make censoring age the age at menopause
                                      ifelse(fu.meno.age.rev<=start.age, NA, c.age))))

# how many non-subcohort members are excluded from premenopausal data given they 
# experience menopause before the event
# should I assume they are still pre-menopausal at baseline?
summary(df$c.age.alt)

with(df, table(subcohort, event))

ind = with(df, ifelse(c.age.alt<c.age & subcohort==0, 1, 0))
table(ind)

df$event.alt = with(df, ifelse(c.age.alt < c.age & subcohort==0, NA, # if alternative censor age less than original censor age then no event for people not in subcohort
                               ifelse(c.age.alt<c.age & subcohort==1, 0, event)))
summary(df$event.alt)

df = df[which(df$menop.status==0 & !(is.na(df$event.alt))),]
with(df, table(subcohort, event.alt))
dim(df)

```

```{r, eval=F, include=F}
# test out updated censoring and event times based on menopause status
vars = c("start.age", "c.age", "c.age.alt", 
         "fu.meno.age.rev", "menop.status", "event", 
         "event.alt", "fertn", 
         "subcohort")

cc=df[complete.cases(df[vars]),c(vars, "PSID")]
summary(cc)
dim(cc)

ind2 = with(cc, ifelse(c.age.alt<=start.age, 1, 0))
table(ind2)

ccpm = cc[cc$menop.status==0,]

cch(Surv(start.age, c.age.alt, event.alt) ~ fertn,
             data=ccpm,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=50000)

head(df[df$c.age.alt<df$start.age, vars])

with(df, table(subcohort, event.alt, exclude=NULL))
sum(with(df, table(subcohort, event.alt)))

df = df[which(df$menop.status==0),]
dim(df)
with(df, table(subcohort, event))
with(df, table(subcohort, event.alt))
summary(df[df$subcohort==0,]$c.age.alt)


sapply( df[vars], summary)
cc = df[complete.cases(df[vars]),]
dim(df) - dim(cc)  
with(cc, table(subcohort, event.alt))
ind = with(cc, ifelse(c.age.alt<=start.age, 1, 0))
table(ind)
head(cc[ind==1,vars])

dim(df[complete.cases(df[vars]),])
summary(df[vars])
cch(Surv(start.age, c.age.alt, event.alt) ~ fertn,
             data=df[complete.cases(df[vars]),],
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=50000)

```


## High values of ferritin, > 300 $\mu$g/L

```{r}

df$iron.cov1 = with(df, ifelse( (menop.status==1 & UMN_Iron_Baseline_FERTN>300) | 
                                 (menop.status==0 & UMN_Iron_Baseline_FERTN>200), 1,
                              ifelse(!is.na(UMN_Iron_Baseline_FERTN), 0, NA)))
table(df$iron.cov1, useNA = "ifany")

cox.overload = coxph(Surv(start.age, c.age.alt, event.alt) ~ iron.cov1,
             data=df); cox.overload

```

```{r, results='markup'}
cox.overload
```


### Association between breast cancer HR and lower thresholds


```{r}

# function to output HR with 95% ci =========================
get.ci = function(df){
  
  colnames(df)[which(colnames(df) %in% c("SE"))] = "se.coef"
  colnames(df)[which(colnames(df) %in% c("p"))] = "p.value"
  colnames(df)[which(colnames(df) %in% c("Value"))] = "coef"

  # source: https://stackoverflow.com/questions/50118394/selecting-and-colouring-single-table-cells-with-kableextra-in-r-markdown-cell-sp
  
   with(data.frame(df),  ifelse(p.value<0.05/100, # I'm trying to get rid of bolding so makeing really extreme.
                                paste0("\\textbf{",
                                       formatC(round(exp(coef),2), format="f", flag='0', digits=2),
                                       " (",
                                      formatC(round(exp(coef-1.96*se.coef),2), format="f", digits=2),
                                      ", ",
                                      formatC(round(exp(coef+1.96*se.coef),2), format="f", digits=2),
                                      ")}"),
        paste0(formatC(round(exp(coef),2), format="f", flag='0', digits=2),
                                  " (",
                                  formatC(round(exp(coef-1.96*se.coef),2), format="f", digits=2),
                                  ", ",
                                  formatC(round(exp(coef+1.96*se.coef),2), format="f", digits=2),
                                  ")")))
  }

```

```{r, include=F, eval=F}

with(df, table(UMN_Iron_Baseline_FERTN<=12))
with(df, table(UMN_Iron_Baseline_FERTN<=25))
with(df, table(UMN_Iron_Baseline_FERTN<=30))
with(df, table(UMN_Iron_Baseline_FERTN<=45))
with(df, table(UMN_Iron_Baseline_FERTN<=100))


```


```{r}
# make a function of the previous analyses repeating over covariates.

get.coefs = function(threshold, extreme="low", n.size=50784) {

# threshold=12; extreme="low" ; n.size=50784 # debug
var.fe = "fertn"

# designate the iron covariate  
iron.var = if(extreme == "low") {
  with(df, ifelse(UMN_Iron_Baseline_FERTN<=threshold, 1, 
                               ifelse(!is.na(UMN_Iron_Baseline_FERTN), 0, NA)))
  } else {
    with(df, ifelse(UMN_Iron_Baseline_FERTN>=threshold, 1, 
                               ifelse(!is.na(UMN_Iron_Baseline_FERTN), 0, NA)))
  }

df$iron.var = iron.var
summary(df$iron.var)
df  = df[complete.cases(df$iron.var),]

# 1) Cox ph with quartiles ==================================


# Note: There must not be any censored observations outside the subcohort.
with(df, table(subcohort, event.alt))
cc = df[complete.cases(df[c("start.age", "c.age.alt", "event.alt", "iron.var", "subcohort",
                            "PSID")]),]
dim(df)-dim(cc) 

cox1 = cch(Surv(start.age, c.age.alt, event.alt) ~ iron.var,
             data=cc,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size)

cq = coef(summary(cox1))
coefs.1 = get.ci(cq)[1]
m.event = summary(cox1)$nevent
tot = nrow(df)

n.threshold = table(iron.var)[2] # number < that are in subcohort
n.threshold.sub = table(df$iron.var,df$subcohort)[2,2] # number < that are in subcohort
sub.size = cox1$subcohort.size

cox1.info = data.frame(threshold = threshold,
                       coef = coefs.1,
#                       n.threshold= paste0(n.threshold, " (", tot, ")" )
                      n.threshold=n.threshold)

# put all this info into a list of objects to output from the function =====================

return(list(cox1.info))

}

# get.coefs( threshold=30) # check

```

```{r}

# lower thresholds
thresh = c(12, 25, 45)
  
models1 = lapply(thresh, function(x) get.coefs(x, "low"))

models1

```

```{r}

# upper thresholds
thresh = c(200,300)
  
models.high = lapply(thresh, function(x) get.coefs(x, "high"))

models.high

```


```{r}

# extract out first element from each list, the coefficients to print off to table
coefs.list = lapply(models1, '[[', 1)  # This returns a list with only the third element, the extracted coef data frame
coefs.list


coefs.list.h = lapply(models.high, '[[', 1)  # This returns a list with only the third element, the extracted coef data frame
coefs.list.h

cox.dat.premenop1 =  coefs.list %>% bind_rows() # source: https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function
cox.dat.premenop1$high=0

cox.dat.premenop2 =  coefs.list.h %>% bind_rows() 
cox.dat.premenop2$high=1


cox.dat.pre = rbind(cox.dat.premenop1, cox.dat.premenop2)
factor(cox.dat.pre$threshold)
cox.dat.pre$threshold.f = with(cox.dat.pre, factor(threshold, 
                                         labels = c("$\\leq$ 12",
                                                    "$\\leq$ 25",
                                                    "$\\leq$ 45",
                                                    "$\\geq$ 200",
                                                    "$\\geq$ 300")))
cox.dat.pre

save(cox.dat.pre, file="../sections/k4-pre.RData")

```


```{r, results='markup'}
kable(cox.dat.pre, 
      booktabs=T, 
      col.names = c("Ferritin $\\leq$ threshold ($\\mu$g/L)", "HR (95\\% CI)", "n below", "High/low", 
                    "threshold"),
      escape=F)

```

