---
title: "Section 5: Post-hoc 2: Association between extreme transferrin saturation (TSAT) levels and breast cancer risk (case-cohort analysis). postmenopausal"
output:
  pdf_document:
    toc: no
    number_sections: true
    includes:
      in_header: header.tex 
  html_document:
    theme: united
    toc: yes
geometry: margin=2cm
always_allow_html: true
---


# Section 5. Post-hoc 2: Association between extreme transferrin saturation (TSAT) levels and breast cancer risk (case-cohort analysis)


- High transferrin saturation levels are considered to be > 45%, > 50%, > 55%, and > 60%. All of these cutoff values had previously been proposed or used in population-based studies of elevated serum transferrin saturation [@mainousElevatedSerumTransferrin2004].

- Low transferring saturation levels can be as low as 15% and up to 25% thresholds [@peyrin-birouletGuidelinesDiagnosisTreatment2015]


```{r, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      results = 'markup',
                      warning = F)

```


```{r, include=F}
# bring in packages
require(knitr)
require(ggplot2)
require(kableExtra)
require(survival)
require(ggplot2)
require(dplyr)
```


```{r}

# Source of data: section2-cch-postmenopause.Rmd
load(file="data-post.RData") # df.pc

df = data.frame(df.pc)

# select only cases and subcohort

df = df[df$event==1 | df$subcohort==1,]
dim(df)

```


```{r, results='hide'}

# handle data for survival analysis

df = within(df, {


  fe.log = log(UMN_Iron_Baseline_FE)
  fertn.log = log(UMN_Iron_Baseline_FERTN)
  fesat.log = log(UMN_Iron_Baseline_FESAT)
  tibc.log = log(UMN_Iron_Baseline_TIBC)
  uibc.log = log(UMN_Iron_Baseline_UIBC)
    
  fe.scale = scale(log(UMN_Iron_Baseline_FE), scale=F)
  fertn.scale = scale(log(UMN_Iron_Baseline_FERTN), scale=F)
  fesat.scale = scale(log(UMN_Iron_Baseline_FESAT), scale=F)
  tibc.scale = scale(log(UMN_Iron_Baseline_TIBC), scale=F)
  uibc.scale = scale(log(UMN_Iron_Baseline_UIBC), scale=F)

})

confounder.list = c('smoke.f2', 'alc.f2', 'educ.f2', 'ever.hrt.f', 'birth.control.f',
                    'one.term.birth.f')


# time-dept menopause ==============================

summary(df$fu.meno.age) # 892 missing
df$miss.meno.age = with(df, is.na(df$fu.meno.age))
with(df, table(miss.meno.age, menop.status)) 

# should I assume they are still pre-menopausal at baseline?
df$fu.meno.age.rev = with(df, ifelse( is.na(fu.meno.age)==T & menop.status==1, baseline.age,
                                    ifelse(is.na(fu.meno.age)==T & menop.status==0, 99, fu.meno.age )))
summary(df$fu.meno.age.rev)

df[df$PSID %in% c("00224_100003"), c("fu.meno.age", "baseline.age", "menop.status", "fu.meno.age.rev")]

df$miss.meno.age.rev = with(df, is.na(df$fu.meno.age.rev))
with(df, table(miss.meno.age.rev, menop.status)) # check

```

## FE Saturation

```{r}

df$iron.cov2 = with(df, ifelse(UMN_Iron_Baseline_FESAT<12, 1, 
                               ifelse(!is.na(UMN_Iron_Baseline_FESAT), 0, NA)))
table(df$iron.cov2, useNA = "ifany")

cox.deficiency = coxph(Surv(start.age, c.age, event) ~ iron.cov2,
             data=df); cox.deficiency
```


```{r}

# function to output HR with 95% ci =========================
get.ci = function(df){
  
  colnames(df)[which(colnames(df) %in% c("SE"))] = "se.coef"
  colnames(df)[which(colnames(df) %in% c("p"))] = "p.value"
  colnames(df)[which(colnames(df) %in% c("Value"))] = "coef"

  # source: https://stackoverflow.com/questions/50118394/selecting-and-colouring-single-table-cells-with-kableextra-in-r-markdown-cell-sp
  
   with(data.frame(df),  ifelse(p.value<0.05/10,
                                paste0("\\textbf{",
                                       formatC(round(exp(coef),2), format="f", flag='0', digits=2),
                                       " (",
                                      formatC(round(exp(coef-1.96*se.coef),2), format="f", digits=2),
                                      ", ",
                                      formatC(round(exp(coef+1.96*se.coef),2), format="f", digits=2),
                                      ")}"),
        paste0(formatC(round(exp(coef),2), format="f", flag='0', digits=2),
                                  " (",
                                  formatC(round(exp(coef-1.96*se.coef),2), format="f", digits=2),
                                  ", ",
                                  formatC(round(exp(coef+1.96*se.coef),2), format="f", digits=2),
                                  ")")))
  }

```


```{r}
# make a function of the previous analyses repeating over covariates.

get.coefs = function(threshold, extreme="low", n.size=50784) {

var.fe = "fesat"
# threshold=30; extreme="low" ; n.size=50784 # debug

# designate the iron covariate  
iron.var = if(extreme == "low") {
  with(df, ifelse(UMN_Iron_Baseline_FESAT<=threshold, 1, 
                               ifelse(!is.na(UMN_Iron_Baseline_FERTN), 0, NA)))
  } else {
    with(df, ifelse(UMN_Iron_Baseline_FESAT>=threshold, 1, 
                               ifelse(!is.na(UMN_Iron_Baseline_FERTN), 0, NA)))
  }

df$iron.var = iron.var
summary(df$iron.var)
df  = df[complete.cases(df$iron.var),]



# 1) Cox ph with quartiles ==================================

cox1 = cch(Surv(start.age, c.age, event) ~ iron.var ,
             data=df,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size)

cox1 = cch(Surv(start.age, c.age, event) ~ iron.var ,
             data=df,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size) 

names(summary(cox1))

cq = coef(summary(cox1))
coefs.1 = get.ci(cq)[1]
m.event = summary(cox1)$nevent
tot = nrow(df)

n.threshold = table(iron.var)[2] # number < that are in subcohort
n.threshold.sub = table(df$iron.var,df$subcohort)[2,2] # number < that are in subcohort
sub.size = cox1$subcohort.size

cox1.info = data.frame(threshold = threshold,
                      coef = coefs.1,
                      n.threshold = n.threshold)
#                      n.threshold= paste0(n.threshold, " (", tot, ")" ))

# put all this info into a list of objects to output from the function =====================

return(list(cox1.info))

}

# get.coefs( threshold=30) # check

```

```{r}

# lower thresholds
thresh = c(20)
  
models1 = lapply(thresh, function(x) get.coefs(x, "low"))

models1

# upper thresholds
thresh2 = c(45, 50, 55, 60)
  
models2 = lapply(thresh2, function(x) get.coefs(x, "high"))

models2


```

## Association between breast cancer HR and lower thresholds

```{r}

# extract out first element from each list, the coefficients to print off to table
coefs.list.1 = lapply(models1, '[[', 1)  # This returns a list with only the third element, the extracted coef data frame
coefs.list.1

# extract out first element from each list, the coefficients to print off to table
coefs.list.2 = lapply(models2, '[[', 1)  # This returns a list with only the third element, the extracted coef data frame
coefs.list.2

cox.dat1 =  coefs.list.1 %>% bind_rows() # source: https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function
cox.dat1
cox.dat1$high=0

cox.dat2 =  coefs.list.2 %>% bind_rows() # source: https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function
cox.dat2
cox.dat2$high=1

cox.dat = rbind(cox.dat1, cox.dat2)
cox.dat2.post = cox.dat

cox.dat2.post$threshold.f = with(cox.dat2.post, factor(threshold, 
                                         labels = c("$\\leq$ 20",
                                                    "$\\geq$ 45",
                                                    "$\\geq$ 50",
                                                    "$\\geq$ 55",
                                                    "$\\geq$ 60" )))
cox.dat2.post


```

```{r, results='markup'}
kable(cox.dat2.post, 
      booktabs=T, 
      col.names = c("TSAT $\\leq$ threshold (\\%)", "HR (95\\% CI)", "n, threshold", "High/low",
                    "threshold"),
      escape=F)


save(cox.dat2.post, file="../sections/k5-post.RData")

```


### Association between breast cancer HR and upper thresholds

```{r}

# extract out first element from each list, the coefficients to print off to table
coefs.list2 = lapply(models2, '[[', 1)  # This returns a list with only the third element, the extracted coef data frame
coefs.list2


cox.dat2.post =  coefs.list2 %>% bind_rows() # source: https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function

cox.dat2.post


```


```{r, results='markup'}
kable(cox.dat2.post, 
      booktabs=T, 
      caption="postmenopausal",
      col.names = c("TSAT $\\geq$ threshold (\\%)", "HR (95\\% CI)", "n, threshold"),
      escape=F)
```

