---
title: "Post-hoc 7: Stratify the primary analyses by racial ethnic groups and BMI (case-cohort analysis)"
output:
  pdf_document:
    toc: yes
    number_sections: true
    includes:
      in_header: header.tex 
  html_document:
    theme: united
    toc: yes
geometry: margin=1.5cm
editor_options:
  chunk_output_type: console
always_allow_html: true
---

# Table 2. Association between iron status and breast cancer risk (case-cohort analysis)

```{r , include=FALSE}
knitr::opts_chunk$set(echo = T, 
                      results = 'markup',
                      warning = F,
                      fig.width=10,
                      fig.height=8)

```


```{r, include=F}
# bring in packages
require(knitr)
require(data.table)
require(tableone)
require(ggplot2)
require(kableExtra)
require(rowr)
require(survival)
require(stargazer)
require(splines)
require(ggplot2)
require(survminer)
require(dplyr)
require(reshape2)
require(ggpubr)
require(tools)
require(Hmisc)
require(janitor)
require(foreign)
require(haven)

```


```{r}

# source: section1.Rmd
load(file="../sections/updated-data-iron.RData") # df3 is the iron subset
names(df3)

df=data.frame(df3[df3$subcohort %in% c(0,1),])
dim(df)
with(df[df$SE_RACE15==5,], table(subcohort, event))

table(as_factor(df$SE_RACE15))
table(as_factor(df$SE_RACE_ETH))

```


```{r}

# handle data for survival analysis
df = within(df, {

  fe.log = log(UMN_Iron_Baseline_FE)
  fertn.log = log(UMN_Iron_Baseline_FERTN)
  fesat.log = log(UMN_Iron_Baseline_FESAT)
  tibc.log = log(UMN_Iron_Baseline_TIBC)
  uibc.log = log(UMN_Iron_Baseline_UIBC)
    
  fe.scale = scale(log(UMN_Iron_Baseline_FE), scale=F)
  fertn.scale = scale(log(UMN_Iron_Baseline_FERTN), scale=F)
  fesat.scale = scale(log(UMN_Iron_Baseline_FESAT), scale=F)
  tibc.scale = scale(log(UMN_Iron_Baseline_TIBC), scale=F)
  uibc.scale = scale(log(UMN_Iron_Baseline_UIBC), scale=F)
})

summary(df$futime)
summary(df$c.age)
summary(df$FU_BCInvD_EOFAge)
summary(df$FU_BCInvD_EOFAgeExact)
sapply(df[c('alc.f3', 'educ.f2', 'ever.hrt.f', 'birth.control.f',
                     'age.menarche', 'age.firstbirth.cat')], summary)
                     
confounder.list = c('alc.f3',  'educ.f2', 'bmi', 'menop.status', 'menop.status:bmi', 
                     'age.menarche', 'age.firstbirth.cat')

outcome.vars = c("fe", "fertn", "fesat")

# time-dept menopause ==============================

summary(df$fu.meno.age) 
df$miss.meno.age = with(df, is.na(df$fu.meno.age))
with(df, table(miss.meno.age, menop.status)) # people are post-menopausal at baseline but no age at menopause. Put age at menopause at baseline age.

# take only non-missing iron values
dim(df)
df =  df[complete.cases(df[outcome.vars]),]
dim(df)

# should I assume they are still pre-menopausal at baseline?
df$fu.meno.age.rev = with(df, ifelse( is.na(fu.meno.age)==T & menop.status==1, baseline.age,
                                    ifelse(is.na(fu.meno.age)==T & menop.status==0, 99, fu.meno.age )))
summary(df$fu.meno.age.rev)

df$miss.meno.age.rev = with(df, is.na(df$fu.meno.age.rev))
with(df, table(miss.meno.age.rev, menop.status)) # check

dim(df)


```


```{r}
# PCA code

# Source: https://raw.githubusercontent.com/WinVector/Examples/master/PCR/XonlyPCA.Rmd
extractProjection <- function(ndim,princ) {
  # pull off the rotation.  
  proj <- princ$rotation[,1:ndim] 
  # sign was arbitrary, so flip in convenient form
  for(i in seq_len(ndim)) {
    si <- sign(mean(proj[,i]))
    if(si!=0) {
      proj[,i] <- proj[,i]*si
    }
  }
  proj
}

```

```{r}
# get first PC

iron.vars = c("fe", 'fesat', "fertn")
iron.log.vars = c("fe.log", "fesat.log", "fertn.log")

# Source: http://www.win-vector.com/blog/2016/05/pcr_part1_xonly/
m.df = as.matrix(df[iron.log.vars]) # use complete cases

pca.log = prcomp(m.df, center=T, scale.=T) # scale variables because units are not same across 3 measures.

# signs are arbitrary on PCA, so instead of calling predict we pull out
# (and alter) the projection by hand

projectedTrainIdeal <-
  as.data.frame(scale(m.df) %*% extractProjection(3,pca.log),
                                 stringsAsFactors = FALSE)
head(projectedTrainIdeal)
head(m.df)


df.pc = cbind(projectedTrainIdeal, df)
with(df.pc, cor(PC1, fe))

```


```{r}

# function to output HR with 95% ci =========================
get.ci = function(df){
  
  colnames(df)[which(colnames(df) %in% c("SE"))] = "se.coef"
  colnames(df)[which(colnames(df) %in% c("p"))] = "p.value"
  colnames(df)[which(colnames(df) %in% c("Value"))] = "coef"

  
with(data.frame(df), paste0(formatC(round(exp(coef),2), format="f", flag='0', digits=2),
                                  " (",
                                  formatC(round(exp(coef-1.96*se.coef),2), format="f", digits=2),
                                  ", ",
                                  formatC(round(exp(coef+1.96*se.coef),2), format="f", digits=2),
                                  ")")
)
}


```


```{r}

# make a function of the previous analyses repeating over covariates.
table(df$SE_RACE_ETH)
table(as_factor(df$EX_BMI_CDC_final))
#table(df.pc$bmi2)

get.coefs = function(var.fe, n.size=50884, dfsub, sub.label, bmi="no") {

# bmi="no"; var.fe = "fe" ; n.size=50884 ; dfsub = df[df$SE_RACE_ETH==1,]; sub.label = 'white'  # debug
#dfsub=df2[!(is.na(df2$bmi2)) & df2$bmi2=="2) 18.5-24.9",]
  
# bmi="yes"; var.fe = "fe"; dfsub=df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="4) 30+",]; n.size=50884; sub.label="obese"
  
# bmi="yessmall"; var.fe = "PC1"; dfsub=df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="4) 30+",]; n.size=50884; sub.label="obese" # NOTE: the bmi2 variable is created below this function.

# bmi="yes"; var.fe = "fertn.log"; dfsub=df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="4) 30+",]; n.size=50884; sub.label="obese"

df1 = dfsub
dim(dfsub)

# designate the iron covariate  
df1$iron.cov = df1[,var.fe] # source: https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function

df1=df1[which(!(is.na(df1$iron.cov))),] # need to remove missing for cch to work
summary(df1$iron.cov)

# get quartiles
qrts = quantile(df1$iron.cov, c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
qrts
qrts2 = quantile(exp(df1$iron.cov), c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
#qrts = quantile(df1$iron.cov, c(0, 0.1, 0.5, 0.9, 1), na.rm=T)
qrts2

# make a variable with cut points based on quartiles above
df1$fe.cutq = cut(df1$iron.cov, qrts, include.lowest = T, dig.lab=10)
df1$fe.cutq2 = cut(exp(df1$iron.cov), qrts2, include.lowest = T, dig.lab=4)
df1$fe.cutq.round = cut(df1$iron.cov, round(qrts,1), include.lowest = T, dig.lab=10)
table(df1$fe.cutq.round)
table(df1$fe.cutq2)

# what is the median within those intervals? =====================

dt = data.table(df1)
setkey(dt, fe.cutq)
summary(dt$iron.cov)
summary(df1$iron.cov)

medians = dt[,list(median=median(iron.cov, na.rm=T)), by=fe.cutq]
medians = data.frame(medians[complete.cases(medians)])
medians

# rounded medians =====================
setkey(dt, fe.cutq.round)
medians.round = dt[,list(median=median(iron.cov, na.rm=T)), by=fe.cutq.round]
medians.round = data.frame(medians.round[complete.cases(medians.round)])
medians.round

# get median of untransformed serum iron marker
medians.orig = if(var.fe %in% c("fertn.log")) {
  setkey(dt, fe.cutq2); dt[,list(median=median(exp(iron.cov), na.rm=T)), by=fe.cutq2]} else {
  setkey(dt, fe.cutq); dt[,list(median=median(iron.cov, na.rm=T)), by=fe.cutq]
  }
medians.orig = data.frame(medians.orig[complete.cases(medians.orig),])

# get untransformed values if log transform ferritin
if(var.fe=="fertn.log") {
  qrts = quantile(df1$fertn.log, c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
  qrts2 = quantile(df1$fertn, c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
  
  df1$fe.cutq = cut(df1$fertn.log, qrts, include.lowest = T, dig.lab=4)
  df1$fe.cutq2 = cut(df1$fertn, qrts2, include.lowest = T, dig.lab=4)
  
  table(is.na(df1$fe.cutq)); table(df1$fe.cutq)
  table(is.na(df1$fe.cutq2)); table(df1$fe.cutq2)
  
  dt.extra = data.table(df1)
  setkey(dt.extra, fe.cutq2)
  medians =  dt.extra[,list(median=median(fertn, na.rm=T)), by=fe.cutq2]
  medians = data.frame(medians[complete.cases(medians)])
} else {
  # get quartiles
qrts = quantile(df1$iron.cov, c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
#qrts = quantile(df1$iron.cov, c(0, 0.1, 0.5, 0.9, 1), na.rm=T)
qrts

# make a variable with cut points based on quartiles above
df1$fe.cutq = cut(df1$iron.cov, qrts, include.lowest = T)#, dig.lab=4)

# what is the median within those intervals? =====================
dt = data.table(df1)
setkey(dt, fe.cutq)

medians = dt[,list(median=median(iron.cov, na.rm=T)), by=fe.cutq]
medians = data.frame(medians[complete.cases(medians)])
medians
}


# 1) Cox ph with quartiles ==================================

table(df1$UMN_Iron_Subcohort)
table(df1$subcohort)
table(df1$event)
summary(df1$fe.cutq)
with(df1, table(subcohort, event))

sapply(df1[c("start.age", "c.age")], summary)
head(df1[is.na(df1$c.age),])

cox.qrt = cch(Surv(start.age, c.age, event) ~ fe.cutq , 
             data=df1,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size)
cox.qrt
cox.qrt$n
cq = coef(summary(cox.qrt))
names(cox.qrt)
cox.qrt$cohort.size
cox.qrt$subcohort.size

sub.label2 = paste0(sub.label, " (n=", cox.qrt$n, ", subcohort n=", cox.qrt$subcohort.size, ")")
sub.label2

coefs.qrts = get.ci(cq)[1:3]
coefs.qrts
names(summary(cox.qrt))
n.event = summary(cox.qrt)$subcohort.size

tbl.qrt = kable(coef(summary(cox.qrt)), booktabs=T, 
                caption=paste0("iron.cov = ", var.fe, ", subgroup = ", sub.label)) %>%
  kable_styling(latex_options = c("HOLD_position"), full_width = F)


# Cox ph with continuous covariate ===========================
# ========================================================================

df1 = within(df1, 
             quartile.cov <- as.integer(cut(iron.cov, quantile(iron.cov, probs=0:4/4),
                                                             include.lowest=T)))

cox.lin = cch(Surv(start.age, c.age, event) ~ quartile.cov,
             data=df1,
             subcoh = ~subcohort,
                  id = ~PSID,
                  cohort.size = n.size)
cox.lin
cc = coef(summary(cox.lin))

# output coef with 95% ci
coefs.c = get.ci(cc)[1]
coefs.c
coefs.c.p = formatC(round(cc[1,4],3), format="f", digits=3)


# output coef with 95% ci
coefs.lin = get.ci(cc)[1]
coefs.lin

tbl.lin = kable(coef(summary(cox.lin)), booktabs=T,
               caption=paste0("iron.cov = ", var.fe, ", subgroup = ", sub.label)) %>%
  kable_styling(latex_options = c("HOLD_position"), full_width = F)


# Adjusted results for quartile analyses
# ==============================================


sapply(df1[c("menop.age", "time.since.menop", "yrs.mens", "early.menop.45", "age.menarche", "parity")], summary) # check
sapply(df1[c("menop.age", "time.since.menop", "yrs.mens", "early.menop.45", "age.menarche", "parity")], class) # check


confounder.list.baseline = c('bmi', "early.menop.45", "bmi:early.menop.45")
list.cc = c('bmi', "early.menop.45")#menop.status.f")

confounder.list.baseline.cc.small = c("early.menop.45")
list.cc.small = c("early.menop.45")

confounder.list.baseline.nobmi = c('alc.f4',  'educ.f2',  'birth.control.f', "ever.hrt.f", 'smoke.f2',
                     'age.menarche', 'age.firstbirth.cat', "early.menop.45")#, "menop.status.f")
list.cc.nobmi = c('alc.f4', 'educ.f2', 'birth.control.f', "ever.hrt.f", 'smoke.f2',
                     'age.menarche', 'age.firstbirth.cat', "early.menop.45")#, "menop.status.f")

# if stratified by BMI then exclude from list of covariates/confounders

summary(df1[complete.cases(df1[list.cc.small]),c("fe.cutq", "early.menop.45")])

if(bmi=="no"){ # if no bmi strata then adjust for smaller list of variables
  form.qrt.adj = as.formula(paste0("Surv(start.age, c.age, event) ~ ", 
                         paste0(c("fe.cutq",
                                  confounder.list.baseline), collapse = " + ")))
  cox.qrt.adj= cch(form.qrt.adj,
                  data=df1[complete.cases(df1[list.cc]),], # take only complete cases
                  subcoh = ~subcohort,
                  id = ~PSID,
                  cohort.size = n.size)
} else if (bmi=="yessmall") {
  form.qrt.adj = as.formula(paste0("Surv(start.age, c.age, event) ~ ", 
                         paste0(c("fe.cutq",
                                  confounder.list.baseline.cc.small), collapse = " + ")))
  cox.qrt.adj= cch(form.qrt.adj,
                  data=df1[complete.cases(df1[list.cc.small]),], # take only complete cases
                  subcoh = ~subcohort,
                  id = ~PSID,
                  cohort.size = n.size)
} else if (bmi=="yes") {
  form.qrt.adj = as.formula(paste0("Surv(start.age, c.age, event) ~ ", 
                         paste0(c("fe.cutq",
                                  confounder.list.baseline.nobmi), collapse = " + ")))
  coxph(form.qrt.adj, data=df1)
  cox.qrt.adj= cch(form.qrt.adj,
                  data=df1[complete.cases(df1[list.cc.nobmi]),], # take only complete cases
                  subcoh = ~subcohort,
                  id = ~PSID,
                  cohort.size = n.size)
}

cox.qrt.adj
cqa = coef(summary(cox.qrt.adj))

# output coef with 95% ci
coefs.qrts.adj = get.ci(cqa)[1:3]
coefs.qrts.adj
n.event.adj = summary(cox.qrt.adj)$nevent

tbl.qrt.adj = kable(coef(summary(cox.qrt.adj)), booktabs=T,
                    caption=paste0("iron.cov = ", var.fe))  %>%
  kable_styling(latex_options = c("HOLD_position"), full_width = F)


# Cox ph with continuous covariate, adjusted =========================
# ==========================================================================

if(bmi=="no"){ # if no bmi strata then adjust for smaller list of variables
  form.lin.adj = as.formula(paste0("Surv(start.age, c.age, event) ~ ", 
                           paste0(c("quartile.cov",
                                    confounder.list.baseline), collapse = " + ")))
  
  cox.lin.adj= cch(form.lin.adj,
                    data=df1[complete.cases(df1[list.cc]),], # take only complete cases
                    subcoh = ~subcohort,
                    id = ~PSID,
                    cohort.size = n.size)
} else if (bmi=="yessmall") {
  form.lin.adj = as.formula(paste0("Surv(start.age, c.age, event) ~ ", 
                           paste0(c("quartile.cov",
                                    confounder.list.baseline.cc.small), collapse = " + ")))
  
  cox.lin.adj= cch(form.lin.adj,
                    data=df1[complete.cases(df1[list.cc.small]),], # take only complete cases
                    subcoh = ~subcohort,
                    id = ~PSID,
                    cohort.size = n.size)

} else if (bmi=="yes") {
  form.lin.adj = as.formula(paste0("Surv(start.age, c.age, event) ~ ", 
                           paste0(c("quartile.cov",
                                    confounder.list.baseline.nobmi), collapse = " + ")))
  
  cox.lin.adj= cch(form.lin.adj,
                    data=df1[complete.cases(df1[list.cc.nobmi]),], # take only complete cases
                    subcoh = ~subcohort,
                    id = ~PSID,
                    cohort.size = n.size)

}


cox.lin.adj
ca = coef(summary(cox.lin.adj))

# output coef with 95% ci
coefs.ca = get.ci(ca)[1]
coefs.ca
coefs.ca.p = formatC(round(ca[1,4],3), format="f", digits=3)

# output coef with 95% ci
coefs.lin.adj = get.ci(ca)[1]
coefs.lin.adj

tbl.lin.adj = kable(coef(summary(cox.lin.adj)), booktabs=T, 
                    caption=paste0("iron.cov = ", var.fe))  %>%
  kable_styling(latex_options = c("HOLD_position"), full_width = F)


# make data frame of results
# ===============================================================

colnames(medians)[1]="fe.cutq" # fix for the log transformed analyses

df.info.adj = data.frame(cov.fe = var.fe,
                      groups = c( "ref", levels(medians.round[,"fe.cutq.round"])[2:4], "cont"),
                      coef =   c( "ref", coefs.qrts.adj, coefs.lin.adj),
                      adj = "Yes",
                      sub = sub.label2)

df.info.unadj = data.frame(cov.fe = var.fe,
                      groups = c("ref", levels(medians.round[,"fe.cutq.round"])[2:4], "cont"),
                      coef = c( "ref", coefs.qrts, coefs.lin), 
                      adj="No",
                      sub = sub.label2)


if(var.fe=="PC1") {
  df.info.ranges = data.frame(cov.fe = var.fe,
                      groups =c("ref", levels(medians[,"fe.cutq"])[2:4], "cont"),
                      coef =  c(levels(medians[,"fe.cutq"])[1:4], NA),
                      adj="Ranges",
                      sub = sub.label2)
  } else {
    df.info.ranges = data.frame(cov.fe = var.fe,
                      groups =c("ref", levels(medians[,"fe.cutq"])[2:4], "cont"),
                      coef =  c(levels(medians[,"fe.cutq"])[1:4], NA),
                      adj="Ranges",
                      sub = sub.label2)
                    }

df.info.median = data.frame(cov.fe = var.fe,
                      groups = c("ref", levels(medians.round[,"fe.cutq.round"])[2:4], "cont"),
                      coef = c(formatC(round(medians[,"median"],0), format="f", digits=0), NA),
                      adj="Median",
                      sub = sub.label2)

df.info  = rbind.data.frame(df.info.median,
                            df.info.unadj, 
                            df.info.adj,
                            df.info.ranges)

# put all this info into a list of objects to output from the function =====================

return(list(df.info, tbl.qrt, tbl.lin))

}

table(as_factor(df$SE_RACE_ETH))
dfsub1 = df[df$SE_RACE_ETH==2,]
get.coefs( var.fe = "UMN_Iron_Baseline_FE", dfsub = dfsub1, sub.label="Hispanic")[[1]] # check
get.coefs( var.fe = "fertn.log", dfsub = dfsub1, sub.label="Hispanic")[[1]] # check

```


```{r}

# Run function over all five covariates
o.vars = c("fe", "fertn.log", "fesat", "PC1")
outcome.vars2 = c("fe.scale", "fertn.scale", "fesat.scale", "tibc.scale", "uibc.scale")

names(df)

outcome.vars3 = c("UMN_Iron_Baseline_FE",
                  "UMN_Iron_Baseline_FERTN",
                  "UMN_Iron_Baseline_FESAT",
                  "UMN_Iron_Baseline_TIBC",
                  "UMN_Iron_Baseline_UIBC")

get.coefs( var.fe = "UMN_Iron_Baseline_FE", dfsub = dfsub1, sub.label="white") # check
get.coefs( var.fe = "PC1", dfsub = df.pc, sub.label="white")[[1]] # check
# get.coefs( var.fe = "fertn.log", dfsub=df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="4) 30+",],
#            sub.label="Obese: 30+")

```

```{r}
# Run each model separately by racial/ethnic groups

table(as_factor(df$SE_RACE_ETH))

race.dat = mapply(get.coefs, 
       o.vars,
       dfsub = rep(list(df.pc[df.pc$SE_RACE_ETH==0,],
                        df.pc[df.pc$SE_RACE_ETH==1,],
                        df.pc[df.pc$SE_RACE_ETH==2,]),
                   each=4), # repeat for each of 3 biomarkers
       sub.label = rep(c("White, non-Hispanic",
                         "Black, non-Hispanic",
                         "Hispanic"), each=4),
       bmi="no")

l.tot = length(race.dat);

race.tables = race.dat[ seq(1, l.tot, by=3)] 
race.models = race.dat[seq(2,l.tot, by=3)] # even nums
race.models.lin = race.dat[seq(3,l.tot, by=3)] # even nums

# combine all tables together
race.df = race.tables %>% bind_rows()
race.df

```


```{r}

levels(factor(race.df$adj))
levels(factor(race.df$cov.fe))
levels(factor(race.df$sub))

race.df$cov.fe = factor(race.df$cov.fe,
                           levels=c("fe", "fertn.log", "fesat", "PC1"),
                           labels=c("Iron ($\\mu$g/dL)", 
                                    "Ferritin$^b$ ($\\mu$g/dL)", 
                                    "Transferrin Saturation (\\%)", 
                                    "First principal component"))

race.df$adj = factor(race.df$adj, levels=c("Median", "Ranges", "No", "Yes"),
                     labels=c("Median", "Ranges", "Unadjusted HR (95\\% CI)", "Adjusted$^a$ HR (95\\% CI)"))
head(race.df)
table(race.df$groups)
race.df$quart = rep(c("ref", "Q2", "Q3", "Q4", "linear"), nrow(race.df)/5)
head(race.df)

race.df.wide = dcast(race.df, cov.fe + sub + adj ~ quart, value.var="coef")
race.df.wide

```


```{r}
# Run each model separately by BMI groups

bmi.names = names(table(as_factor(df$EX_BMI_CDC_final)))
with(df, table(EX_BMI_CDC_final))
summary(df$EX_BMI_CDC_final)

# combine all obese into one category
df.pc$bmi2 = factor(df.pc$EX_BMI_CDC_final,
                  labels = c(bmi.names[1:3], rep("4) 30+", 3)))
table(df.pc$bmi2)
summary(df.pc$bmi2)
levels(df.pc$bmi2)

summary(df.pc[complete.cases(df.pc$bmi2) & df.pc$bmi2=="2) 18.5-24.9",]$bmi2)
summary(df.pc[complete.cases(df.pc$bmi2) & df.pc$bmi2=="2) 18.5-24.9",]$subcohort)
summary(df.pc[df.pc$EX_BMI_CDC_final==2,]$bmi2)

# run function over all outcome and exposure variables
bmi.dat = mapply(get.coefs, 
       o.vars,
       dfsub = rep(list(df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="2) 18.5-24.9",],
                        df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="3) 25.0-29.9",],
                        df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="4) 30+",]),
                   each=4), # repeat for each of 4 biomarkers
       sub.label = rep(c("Normal: 18.5-24.9", "Overweight: 25.0-29.9", "Obese: 30+"), each=4),
       bmi="yes")

dim(bmi.dat)

l.totb = length(bmi.dat); l.totb

bmi.tables = bmi.dat[ seq(1, l.totb, by=3)] 
bmi.models = bmi.dat[seq(2,l.totb, by=3)] 
bmi.models.lin = bmi.dat[seq(3,l.totb, by=3)] 


# combine all tables together
bmi.dat.df = bmi.tables %>% bind_rows()
bmi.dat.df

 bmi.dat.df$sub = factor(bmi.dat.df$sub,
                         levels = c("Normal: 18.5-24.9 (n=2094, subcohort n=1105)",
                                    "Overweight: 25.0-29.9 (n=1839, subcohort n=942)",
                                    "Obese: 30+ (n=1760, subcohort n=889)")) # NOTE: to revise order, must have the exact text, including the sample size

```




```{r}

levels(factor(bmi.dat.df$adj))
levels(factor(bmi.dat.df$cov.fe))
levels(factor(bmi.dat.df$sub))

bmi.dat.df$cov.fe = factor(bmi.dat.df$cov.fe,
                           levels=c("fe", "fertn.log", "fesat", "PC1"),
                           labels=c("Iron ($\\mu$g/dL)", 
                                    "Ferritin$^b$ ($\\mu$g/dL)", 
                                    "Transferrin Saturation (\\%)",
                                    "First principal component"))
names(bmi.dat.df)


# bmi.dat.df$sub = factor(bmi.dat.df$sub,
#                         levels = c("Normal: 18.5-24.9", "Overweight: 25.0-29.9", "Obese: 30+"))


bmi.dat.df$adj = factor(bmi.dat.df$adj, levels=c("Median", "Ranges", "No", "Yes"),
                     labels=c("Median", "Ranges", "Unadjusted HR (95\\% CI)", "Adjusted$^a$ HR (95\\% CI)"))
head(bmi.dat.df)
bmi.dat.df$quart = rep(c("ref", "Q2", "Q3", "Q4", "linear"), nrow(bmi.dat.df)/5)
head(bmi.dat.df)

bmi.dat.df.wide = dcast(bmi.dat.df, cov.fe + sub + adj ~ quart, value.var="coef")
bmi.dat.df.wide

```




```{r}

# Run each model separately by BMI groups
# FOR THIS ITERATION, use smaller subset of confounders to see if difference.

# run function over all outcome and exposure variables (only menopause)
bmi.dat2 = mapply(get.coefs, 
       o.vars,
       dfsub = rep(list(df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="2) 18.5-24.9",],
                        df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="3) 25.0-29.9",],
                        df.pc[!(is.na(df.pc$bmi2)) & df.pc$bmi2=="4) 30+",]),
                   each=4), # repeat for each of 4 biomarkers
       sub.label = rep(c("Normal: 18.5-24.9", "Overweight: 25.0-29.9", "Obese: 30+"), each=4),
       bmi="yessmall")


l.totb = length(bmi.dat2); l.totb

bmi.tables2 = bmi.dat2[ seq(1, l.totb, by=3)] 
bmi.models2 = bmi.dat2[seq(2, l.totb, by=3)] 
bmi.models.lin2 = bmi.dat2[seq(3, l.totb, by=3)] 


# combine all tables together
bmi.dat.df2 = bmi.tables2 %>% bind_rows()
bmi.dat.df2


```




```{r}

levels(factor(bmi.dat.df2$adj))
levels(factor(bmi.dat.df2$cov.fe))
levels(factor(bmi.dat.df2$sub))

bmi.dat.df2$cov.fe = factor(bmi.dat.df2$cov.fe,
                           levels=c("fe", "fertn.log", "fesat", "PC1"),
                           labels=c("Iron ($\\mu$g/dL)", 
                                    "Ferritin$^b$ ($\\mu$g/dL)", 
                                    "Transferrin Saturation (\\%)",
                                    "First principal component"))

 bmi.dat.df2$sub = factor(bmi.dat.df2$sub,
                         levels = c("Normal: 18.5-24.9 (n=2094, subcohort n=1105)",
                                    "Overweight: 25.0-29.9 (n=1839, subcohort n=942)",
                                    "Obese: 30+ (n=1760, subcohort n=889)")) # NOTE: to revise order, must have the exact text, including the sample size


bmi.dat.df2$adj = factor(bmi.dat.df2$adj, levels=c("Median", "Ranges", "No", "Yes"),
                     labels=c("Median", "Ranges", "Unadjusted HR (95\\% CI)", "Adjusted$^a$ HR (95\\% CI)"))

tail(bmi.dat.df2,100)

head(bmi.dat.df2)
bmi.dat.df2$quart = rep(c("ref", "Q2", "Q3", "Q4", "linear"), nrow(bmi.dat.df2)/5)
head(bmi.dat.df2)


bmi.dat.df.wide2 = dcast(bmi.dat.df2, cov.fe + sub + adj ~ quart, value.var="coef")
bmi.dat.df.wide2

```



## Summary tables

### Race/ethnicity

```{r}
# print off table

race.df.wide[race.df.wide=="NA"] <- "" # get rid of na
race.df.wide[is.na(race.df.wide)] <- "" # get rid of nadim(race.df.wide)

names.race= names(race.df.wide) 
race.df.wide2 = race.df.wide[c(names.race[1:3], "ref", "Q2", "Q3", "Q4", "linear")]

k2.race = kable(race.df.wide2,
     align='lllllllc',
      caption = "Breast cancer hazard ratios by serum iron covariate and race/ethnic subgroups.",
      booktabs=T,
      col.names = c( "Iron",
                     "Race/ethnicity",
                     "Values", 
                     "ref", "Q2", "Q3", "Q4", ""),
      #align=rep("l", 8),
      escape=F) %>%
  column_spec(c(2), width = "3cm") %>% # NOTE: this has to come first to change row width?
  column_spec(c(8), width = "4cm") %>% # NOTE: this has to come first to change row width?
  collapse_rows(columns = 1:2, valign = "middle") %>% # note: valign no longer works
  add_header_above(c("", "", "", "Quartile of iron-related biomarker" = 4, "Linear trend over quartiles$^c$" = 1), escape=F) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10) %>%
  footnote(alphabet=c("Adjusted for BMI, baseline menopause status, and a product term of these two variables. Number of adjusted factors is small due to small sample size for some groups.",
                      "log transformed",
                      "Iron covariate in quartile units"),
           threeparttable=T) # source: https://rdrr.io/cran/kableExtra/man/footnote.html)

save(k2.race, file="../sections/k2-race.RData")

```



```{r, results='markup', eval=T}
k2.race
```

### BMI

```{r, results='markup'}
# print off table

bmi.dat.df.wide[bmi.dat.df.wide=="NA"] <- "" # get rid of na
bmi.dat.df.wide[is.na(bmi.dat.df.wide)] <- "" # get rid of na
dim(bmi.dat.df.wide)

names.bmi = names(bmi.dat.df.wide) 
bmi.dat.df.wide.tab = bmi.dat.df.wide[c(names.bmi[1:3], "ref", "Q2", "Q3", "Q4", "linear")]

k2.bmi = kable(bmi.dat.df.wide.tab,
               align='lllllllc',
      caption = "Breast cancer hazard ratios by serum iron covariate and BMI subgroups",
      booktabs=T,
      col.names = c( "Iron",
                     "BMI groups",
                     "Values", 
                     "ref$^c$", "Q2", "Q3", "Q4", ""),
      escape=F) %>%
  column_spec(c(2), width = "3cm") %>% # NOTE: this has to come first to change row width?
  column_spec(c(8), width = "4cm") %>% # NOTE: this has to come first to change row width?
  collapse_rows(columns = 1:2, valign = "middle") %>% # note: valign no longer works
  add_header_above(c("", "", "", "Quartile of iron-related biomarker" = 4, "Linear trend over quartiles$^d$" = 1), escape=F) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10)  %>%
    footnote(alphabet = c("Adjusted for baseline smoking, alcohol, education, HRT, age at menarche, age at first birth, oral contraceptive use, and menopause status.",
                          "log transformed",
                          "For first quartile compared to aggregate of upper three quartiles, the adjusted HR (95\\\\% CI) is 0.72 (0.56, 0.92) for iron and 0.84 (0.66, 1.09) for ferritin among women with a BMI $\\\\geq$ 30 kg/m$^2$.",
                          "Iron covariate in quartile units"), 
             threeparttable=T, escape=F)  # source: https://rdrr.io/cran/kableExtra/man/footnote.html


save(k2.bmi, bmi.dat.df.wide.tab, file="../sections/k2-bmi.RData")

```

```{r, results='markup'}
k2.bmi
```

#### Look at ferritin for one subgroup (obese) from the BMI analyses

```{r, results='hide'}

# 3/31/2020 Additional analysis estimating breast cancer HR for lowest quartile vs 2nd - 4th quartile
# adapted from get.coef function
var.fe = "fertn"
df.pc$var.fe = df.pc[var.fe]
summary(df.pc$var.fe); summary(df.pc$fertn)

df.e = df.pc[!(is.na(df.pc$fertn)),] # need to remove missing for cch to work
summary(df.e$fertn)

df.e2 = df.e[!(is.na(df.e$bmi2)) & df.e$bmi2=="4) 30+",]

table(df.e2$UMN_Iron_Subcohort)
table(df.e2$subcohort)
table(df.e2$event)
with(df.e2, table(subcohort, event))

# get quartiles
qrts = quantile(df.e2$fertn, c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
qrts2 = quantile(df.e2$fertn, c(0, 0.25, 1), na.rm=T)
#qrts = quantile(df1$iron.cov, c(0, 0.1, 0.5, 0.9, 1), na.rm=T)
qrts; qrts2

# Make a variable with cut points based on quartiles above
df.e2$fe.cutq = cut(df.e2$fertn, qrts, include.lowest = T, dig.lab=10)
summary(df.e2$fe.cutq)
df.e2$fe.cutq2 = cut(df.e2$fertn, qrts2, include.lowest = T, dig.lab=10)
summary(df.e2$fe.cutq2)

# What is the median within those intervals? =====================
dt = data.table(df.e2)
setkey(dt, fe.cutq)
medians = dt[,list(median=median(fertn, na.rm=T)), by=fe.cutq]
medians = data.frame(medians[complete.cases(medians)])
medians

# get median of untransformed serum iron marker
medians.orig = if(var.fe %in% c("fertn.log")) dt[,list(median=median(exp(fertn.log), na.rm=T)), by=fe.cutq] else dt[,list(median=median(fertn, na.rm=T)), by=fe.cutq]
medians.orig = data.frame(medians.orig[complete.cases(medians.orig)])
medians.orig

# 1) Cox ph with quartiles ==================================

sapply(df.e2[c("start.age", "c.age")], summary)

n.size=50884

# original analysis
cox.qrt = cch(Surv(start.age, c.age, event) ~ fe.cutq ,
             data=df.e2,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size)

cox.qrt
exp(coef(cox.qrt))
```

```{r, results='markup'}
summary(df.e2$fe.cutq2)
df.e2$fe.cutq2 = relevel(df.e2$fe.cutq2, ref="(40,1404]")

# additional analysis comparing lowest quartile to rest of group
cox.qrt2 = cch(Surv(start.age, c.age, event) ~ fe.cutq2 , # + cluster(HH_PSID) # NOTE: clustering does not work
             data=df.e2,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size)

cox.qrt2
exp(coef(cox.qrt2))

get.ci(coef(summary(cox.qrt2)))

```

#### Look at iron for one subgroup (obese) from the BMI analyses

```{r, results='hide'}

# 3/31/2020 Additional analysis estimating breast cancer HR for lowest quartile vs 2nd - 4th quartile
# adapted from get.coef function
var.fe = "fe"; var.fe
df.pc$var.fe = df.pc[,var.fe]
class(df.pc$var.fe)
summary(df.pc$var.fe); summary(df.pc$fe)

df.e = df.pc[!(is.na(df.pc$var.fe)),] # need to remove missing for cch to work
summary(df.e$var.fe)

df.e2 = df.e[!(is.na(df.e$bmi2)) & df.e$bmi2=="4) 30+",]

# get quartiles
qrts = quantile(df.e2$var.fe, c(0, 0.25, 0.5, 0.75, 1), na.rm=T)
qrts2 = quantile(df.e2$var.fe, c(0, 0.25, 1), na.rm=T)
#qrts = quantile(df1$iron.cov, c(0, 0.1, 0.5, 0.9, 1), na.rm=T)
qrts; qrts2

# Make a variable with cut points based on quartiles above
df.e2$fe.cutq = cut(df.e2$var.fe, qrts, include.lowest = T, dig.lab=10)
summary(df.e2$fe.cutq)
df.e2$fe.cutq2 = cut(df.e2$var.fe, qrts2, include.lowest = T, dig.lab=10)
summary(df.e2$fe.cutq2)

# What is the median within those intervals? =====================
dt = data.table(df.e2)
setkey(dt, fe.cutq)
medians = dt[,list(median=median(fertn, na.rm=T)), by=fe.cutq]
medians = data.frame(medians[complete.cases(medians)])
medians

# get median of untransformed serum iron marker
medians.orig = if(var.fe %in% c("fertn.log")) dt[,list(median=median(exp(fertn.log), na.rm=T)), by=fe.cutq] else dt[,list(median=median(fertn, na.rm=T)), by=fe.cutq]
medians.orig = data.frame(medians.orig[complete.cases(medians.orig)])
medians.orig

```

```{r, results='markup'}
summary(df.e2$fe.cutq2)
df.e2$fe.cutq2 = relevel(df.e2$fe.cutq2, ref="(69,276]")

# additional analysis comparing lowest quartile to rest of group
cox.qrt2 = cch(Surv(start.age, c.age, event) ~ fe.cutq2  , 
             data=df.e2,
             subcoh = ~subcohort,
             id = ~ PSID,
             cohort.size=n.size)

cox.qrt2
exp(coef(cox.qrt2))

get.ci(coef(summary(cox.qrt2)))

```


### BMI, only adjusted for menopause

```{r, results='markup'}

# print off table

bmi.dat.df.wide2[bmi.dat.df.wide2=="NA"] <- "" # get rid of na
bmi.dat.df.wide2[is.na(bmi.dat.df.wide2)] <- "" # get rid of nadim(bmi.dat.df.wide)

names.bmi2 = names(bmi.dat.df.wide2) 
bmi.dat.df.wide.tab2 = bmi.dat.df.wide2[c(names.bmi2[1:3], "ref", "Q2", "Q3", "Q4", "linear")]

k2.bmi2 = kable(bmi.dat.df.wide.tab2,
      caption = "Breast cancer hazard ratios by serum iron covariate and BMI subgroups",
      booktabs=T,
      col.names = c( "Iron",
                     "BMI groups",
                     "Values", 
                     "ref", "Q2", "Q3", "Q4", "Continuous$^b$"),
      escape=F) %>%
  collapse_rows(columns = 1:2, valign = "top") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10)  %>%
    footnote(alphabet = c("Adjusted for baseline menopause status.",
                          "Iron covariate in quartile units"), 
             threeparttable=T) # source: https://rdrr.io/cran/kableExtra/man/footnote.html

save(k2.bmi2, file="../sections/k2-bmi-single.RData")
```

```{r, results='markup'}
k2.bmi2
```


## Unadjusted models for quartiles

### BMI

```{r, results='asis', eval=T}

for(i in bmi.models) print(i) 

```

### Race/ethnicity

```{r, results='asis', eval=T}

for(i in race.models) print(i)

```

## Unadjusted models for continuous variable

### BMI

```{r, results='asis', eval=T}

for(i in bmi.models.lin) print(i) 

```

### Race/ethnicity

```{r, results='asis', eval=T}

for(i in race.models.lin) print(i)

```
