--- 
always_allow_html: yes
bibliography: ../bib/iron-lit.bib
description: Summary of analyses
editor_options:
  chunk_output_type: console
geometry: margin=1.5cm
link-citations: yes
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    fig_caption: yes
    includes:
      in_header: ../preamble-latex-nobookdown.tex
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
    toc: no
---


```{r , include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      results = 'markup',
                      warning = F,
                      fig.width=10,
                      fig.height=8)

```



```{r, echo=F}

load(file="../sections/k1.RData") # section1.Rmd table
load(file="../sections/k2-cch.RData") # section2-cch.Rmd table
load(file="../sections/k2-cch2.RData") # section2-sensitivity-cch2.Rmd

load(file="../sections/k11.RData") # section11.Rmd

load(file="../sections/k4.RData") # cox.dat with Fe extremes from section4-cch.Rmd
load(file="../sections/k4-post.RData") # cox.dat.post with Fe extremes from section4-cch-post.Rmd
load(file="../sections/k4-pre.RData") # cox.dat.pre with Fe extremes from section4-cch-pre.Rmd

load(file="../sections/k5.RData") # cox.dat2 with TSAT extremes from section5-cch.Rmd
load(file="../sections/k5-post.RData") # cox.dat2.post with TSAT extremes from section5-cch-post.Rmd
load(file="../sections/k5-pre.RData") # cox.dat2.pre with TSAT extremes from section5-cch-pre.Rmd

load(file="../sections/k9.RData") # k9 with assn between iron biomarkers and cancer subtype from section9-rev.Rmd

load(file="../sections/k2-race.RData") # k2.race from section10-cch.Rmd
load(file="../sections/k2-bmi.RData") # k2.bmi from section10-cch.Rmd
load(file="../sections/k2-bmi-single.RData") # k2.bmi from section10-cch.Rmd

load(file="../sections/k2-race-pm.RData") # k2.race.pm from section10-cch-postmenoponly.Rmd
load(file="../sections/k2-bmi-pm.RData") # k2.bmi.pm from section10-cch-postmenoponly.Rmd

# extra tables after reviewer comments
# ====================================
load(file="../sections/premenop.RData") # premenop and premenop.table.dat from section2-cch-premenopause.Rmd
load(file="../sections/postmenop.RData") # postmenop and postmenop.table.dat from section2-cch-postmenopause.Rmd

load(file="../sections/tab-stage.RData") # tab.stage, wide.stage., cts4  from reviewer-extra-stage.Rmd

load(file="../sections/k2-bmi-q1.RData") # k2.bmi.q1 from section10-cch-q1.Rmd
load(file="../sections/premenop-q1-ref.RData") # premenop.q1, premenop.table.dat.q1, premenop.table.dat2.q1 from section2-cch-premenopause-q1-ref.Rmd
load(file="../sections/postmenop-q1-ref.RData") # postmenop.q1, postmenop.table.dat.q1 from section2-cch-postmenopause-q1-ref.Rmd

```

```{r, include=F}
require(kableExtra)
```


<div id="refs"></div>

\beginsupplement 


```{r}
k2.cch2
```

```{r}
k11
```


\clearpage
\newpage

```{r}
k2.race
```

\clearpage
\newpage

\vspace{-1em}

<!-- Source: section10-cch.Rmd -->

```{r}
# this table removes the footnote with q1 index group HRs
kable(bmi.dat.df.wide.tab,
               align='lllllllc',
      caption = "Breast cancer hazard ratios by serum iron covariate and BMI subgroups",
      booktabs=T,
      col.names = c( "Iron",
                     "BMI groups",
                     "Values", 
                     "ref", "Q2", "Q3", "Q4", ""),
      escape=F) %>%
  column_spec(c(2), width = "3cm") %>% # NOTE: this has to come first to change row width?
  column_spec(c(8), width = "4cm") %>% # NOTE: this has to come first to change row width?
  collapse_rows(columns = 1:2, valign = "middle") %>% # note: valign no longer works
  add_header_above(c("", "", "", "Quartile of iron-related biomarker" = 4, "Linear trend over quartiles$^c$" = 1), escape=F) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10)  %>%
    footnote(alphabet = c("Adjusted for baseline smoking, alcohol, education, HRT, age at menarche, age at first birth, oral contraceptive use, and menopause status.",
                          "log transformed",
                          "Iron covariate in quartile units"), 
             threeparttable=T, escape=F)  # source: https://rdrr.io/cran/kableExtra/man/footnote.html


```


\vspace{-20em}

```{r}
k2.bmi.q1
```

\clearpage
\newpage

<!-- Source: section2-cch-premenopause.Rmd -->

```{r, results='markup'}
# this table removes the footnote with q1 index group HRs
 kable(premenop.table.dat[,-1],
               align='lllllc',
      caption = "Breast cancer risk associated with increasing levels of iron-related biomarkers for women with premenopausal person-time.",
      booktabs=T,
      col.names = c( "Biomarkers",
                     "Quartile 1$^c$",
                     "Quartile 2", 
                     "Quartile 3",
                     "Quartile 4",
                     " "),
      escape=F) %>%
  #mutate_all(linebreak) %>% # Source: http://haozhu233.github.io/kableExtra/best_practice_for_newline_in_latex_table.pdf
  add_header_above(c(" ",  "Quartile of iron-related biomarker" = 4, "Linear trend over quartiles$^c$" = 1), escape=F) %>%
  pack_rows("Iron ($\\\\mu$g/dL)", 1, 5, escape=F) %>%
  pack_rows("Ferritin$^b$ ($\\\\mu$g/dL)", 6, 10, escape=F) %>%
  pack_rows("Transferrin saturation (\\\\%)", 11, 15, escape=F) %>%
  pack_rows("First principal component", 16, 20) %>%
  column_spec(c(6), width = "4cm") %>% # NOTE: this has to come first to change row width?
  footnote(alphabet = c("Adjusted for  baseline smoking, alcohol, education,  HRT, age at menarche, age at first birth, oral contraceptive use, and BMI.",
                        "log transformed",
                        "Iron covariate in quartile units"), 
           #general="Bold values indicate statistical significance at alpha = 0.05.",
           threeparttable=T) %>% # source: https://rdrr.io/cran/kableExtra/man/footnote.html
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10)

```


\clearpage
\newpage


```{r, results='markup'}
premenop.q1
```


\clearpage
\newpage

```{r, results='markup'}
postmenop
```


\clearpage
\newpage

```{r, results='markup'}
postmenop.q1
```


\clearpage
\newpage



```{r}
cox.dat$fe.val = "Ferritin"
cox.dat2$fe.val = "Transferrin saturation"

ex.dat = rbind(cox.dat[c("fe.val", "threshold", "coef", "n.threshold")],
               cox.dat2[c("fe.val", "threshold", "coef", "n.threshold")])
#ex.dat

```


```{r, include=F, eval=F}
kable(ex.dat[,-1],
      caption="Breast cancer HR by body iron status extreme thresholds",
      booktabs=T,
      col.names = c("Cut points for dichotomized thresholds", "HR (95\\% CI)", "n above/below threshold"),
      escape=F) %>%
  pack_rows("Ferritin $\\\\leq$ threshold ($\\\\mu$g/L)", 1,3, escape=F) %>%
  pack_rows("Transferrin saturation $\\\\geq$ threshold (\\\\%)", 4, 7, escape=F) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size=10)
  
```


```{r}

# cox.dat.post$fe.val = "Ferritin $\\\\leq$ threshold ($\\\\mu$g/L)"
# cox.dat.pre$fe.val = "Ferritin $\\\\leq$ threshold ($\\\\mu$g/L)"
# 
# cox.dat2.post$fe.val = "Transferrin saturation $\\\\geq$ threshold (\\\\%)"
# cox.dat2.pre$fe.val = "Transferrin saturation $\\\\geq$ threshold (\\\\%)"
 
cox.dat.post$fe.val = "Ferritin ($\\mu$g / L)"
cox.dat.pre$fe.val = "Ferritin ($\\mu$g / L)"

cox.dat2.post$fe.val = "Transferrin saturation (\\%)"
cox.dat2.pre$fe.val = "Transferrin saturation (\\%)"

cox.dat.post$menop = "Postmenopausal"
cox.dat2.post$menop = "Postmenopausal"
cox.dat2.pre$menop = "Premenopausal"
cox.dat.pre$menop = "Premenopausal"

ex.dat = rbind(cox.dat.post[c( "fe.val", "menop","threshold.f", "coef", "n.threshold", "high")],
               cox.dat.pre[c( "fe.val", "menop", "threshold.f", "coef", "n.threshold", "high")],
               cox.dat2.post[c( "fe.val", "menop","threshold.f", "coef", "n.threshold", "high")],
               cox.dat2.pre[c( "fe.val", "menop", "threshold.f", "coef", "n.threshold", "high")])
#ex.dat
ex.dat = ex.dat[order(ex.dat$menop, ex.dat$high, ex.dat$fe.val),]

ex.dat$high = factor(ex.dat$high, labels=c("Low", 
                                           "High"))
row.names(ex.dat) = NULL



```


\clearpage
\newpage

```{r}
kable(ex.dat[,c("high", "fe.val", "threshold.f", "coef", "n.threshold")],
      caption="Breast cancer hazard ratios by menopause status at baseline and body iron status extreme thresholds",
      booktabs=T,
      col.names = c("High/low indicator", "Iron measure", 
                    "Cut points for dichotomized iron measure thresholds", "HR (95\\% CI)", 
                    "n above/below threshold"),
      escape=F) %>%
  column_spec(1:4, width = "12em") %>%
  collapse_rows(columns = 1:2, valign = "top") %>%
  pack_rows("Postmenopausal", 1, 10, escape=F) %>%
  pack_rows("Premenopausal", 11, 19, escape=F) %>%
#  pack_rows("Ferritin $\\\\leq$ threshold ($\\\\mu$g/L)", 1,6, escape=F) %>%
#  pack_rows("Transferrin saturation $\\\\geq$ threshold (\\\\%)", 7, 12, escape=F) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10)
  
```


```{r, include=F, eval=F}
kable(ex.dat[,-1],
      caption="Breast cancer hazard ratios by body iron status extreme thresholds",
      booktabs=T,
      col.names = c(  "Menopause status" , 
                    "Cut points for dichotomized iron measure thresholds", "HR (95\\% CI)", "n above/below threshold"),
      escape=F) %>%
  column_spec(1:4, width = "12em") %>%
  collapse_rows(columns = 1, valign = "top") %>%
  pack_rows("Ferritin $\\\\leq$ threshold ($\\\\mu$g/L)", 1,6, escape=F) %>%
  pack_rows("Transferrin saturation $\\\\geq$ threshold (\\\\%)", 7, 12, escape=F) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"), font_size=10)
  
```

\clearpage
\newpage

```{r}

k9

```

\clearpage
\newpage

```{r, results='markup'}
# Odds of breast cancers stages in case-only sample by iron levels
tab.stage
```


\clearpage
\newpage


Supplemental text regarding sample handling and laboratory assessments

The Advanced Research and Diagnostic Laboratory at the University of Minnesota acted as the central laboratory for this study, receiving 7,621 serum samples in straws in a frozen state packaged on dry ice in goblets of 100 samples per goblet. Samples were stored frozen at -80C until testing. To prepare samples for analysis the straw samples were thawed to room temperature, then the straw was cut with a scissor and the sample expelled into a labelled cryovial and capped. Scissors were cleaned with an alcohol pad between each sample. Vials were mixed by inversion immediately before testing. Analysis was performed on the Roche Cobas 6000 Chemistry analyzer (Roche Diagnostics, Mannheim, Germany) using Roche reagents and calibrators. Test methodologies and inter-assay coefficient of variation were as follows. Iron is a colorimetric FerroZine end-point assay with an interassay CV of 2.5\% at a concentration 77 ug/dL. Ferritin is particle enhanced immunoturbidimetric assay with an interassay CV of 2.8\% at a concentration of 259 ug/L. Unsaturated Iron Binding Capacity (UIBC) is a colorimetric assay with an inter-assay CV of 3.3\% at a concentration of 239 ug/dL. All samples had serum indices measurements for hemolysis, icterus and lipemia performed on the Roche Cobas 6000 analyzer by spectrophotometric method. Each assay has manufacture defined thresholds for each indice and results were not reported if this threshold was exceeded. Two levels of quality control were analyzed twice per day, once at the beginning of daily testing and once at the end. One quality control was a laboratory developed pooled serum sample and the other a Roche Precipath U commercially available product. Assay calibration was performed per manufacture specifications using manufacturer calibrators. Sample results were reported to the laboratory information system via electronic instrument interface only after quality control and calibration results were deemed acceptable per laboratory and manufacture standards.

Results that could not be reported were primarily due to sample hemolysis. Our Cobas analyzer is programmed to automatically perform specimen integrity checks for hemolysis, lipemia and icterus. The UIBC assay is particularly sensitive to hemolysis and has the lowest hemolysis index threshold for acceptability which is why more UIBC test results were not reported.